\paragraph{Shape Transition from Mean Pooling to GCN.}
In the original model, we summarized the species dimension using global average
pooling, which collapsed the 115-species alignment into a single feature vector
per position. This approach treats every species equally and discards the
phylogenetic relationships among them.

In GraphyloVar, we replace mean pooling with a two-layer Graph Convolutional
Network (GCN) that operates directly on the phylogenetic adjacency matrix.
The input tensor has shape $(B, 115, F)$ where $B$ is the batch size, 115 is
the number of nodes (58 extant species + 57 inferred ancestors) in the
placental mammalian evolutionary tree, and $F$ is the feature dimension from the
upstream Transformer encoder.

The first GCN layer transforms this to $(B, 115, 32)$ by propagating features
along edges of the phylogenetic graph, with each node aggregating information
from its evolutionary neighbors weighted by the normalized adjacency matrix
$\hat{A} = D^{-1/2} A D^{-1/2}$. The second GCN layer further refines
these representations to $(B, 115, 32)$. The output is then flattened to
$(B, 3680)$ and passed through a shared fully connected layer (128 units, ReLU)
before the dual-output prediction heads (allele frequency distribution and
SNP probability), each with their own 64-unit dense layer.

This architectural change allows the model to learn species-specific weights
that reflect evolutionary importance. \rev{An additive perturbation analysis
(each species evaluated against an all-gap baseline on 500,000 held-out
variants) shows that 53 of 58 extant placental mammals contribute positive
isolated signal, with Human providing the largest contribution ($\Delta$CE
$= 7.1$). Four diverged primates, Orangutan, Green monkey, Gorilla, and
Gibbon, form a secondary cluster ($\Delta$CE $= 0.62$ to $1.50$), and
non-primate mammals contribute a broad gradient ($\Delta$CE $= 0.002$ to $0.12$).
Chimpanzee ranks 11th ($\Delta$CE $= 0.062$) despite being the closest relative
of Human, because its $\approx$98.7\% sequence identity with Human yields
near-zero marginal (leave-one-out) contribution, reflecting information
redundancy with the Human row (see Section~3.6). Five species (Baboon,
Rhesus macaque, Marmoset, Squirrel monkey, Jerboa) show near-zero or marginally
negative additive values, likely due to sparse alignment coverage at the tested
variant sites.} The GCN thus provides an interpretable mechanism for weighting
conservation signals across the phylogenetic tree, rather than assuming uniform
contribution from all species.

